Metric Spaces with Expensive Distances

### 2019-01-24 paper/draft.tex
Correct ANN proof illustration (diameter->radius, centre).

\documentclass[a4paper,USenglish]{socg-lipics-v2018}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathbbol}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{url}
\usepackage{floatflt}
\usepackage{mathptmx}
\usepackage{epsfig}
\usepackage{wrapfig}
\usepackage{color}

@@

\definecolor{uququq}{rgb}{0.25,0.25,0.25}
\definecolor{xdxdff}{rgb}{0.49,0.49,1}
\definecolor{qqqqff}{rgb}{0,0,1}
\begin{figure}[ht!]
    \centering
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=0.5cm,y=0.5cm]
%\draw[->,color=black] (-1.5,0) -- (17,0);
%\foreach \x in {,2,4,6,8,10,12,14,16}
%\draw[shift={(\x,0)},color=black] (0pt,2pt) -- (0pt,-2pt) node[below] {\footnotesize $\x$};
%\draw[->,color=black] (0,-7.5) -- (0,9.2);
%\foreach \y in {-6,-4,-2,2,4,6,8}
%\draw[shift={(0,\y)},color=black] (2pt,0pt) -- (-2pt,0pt) node[left] {\footnotesize $\y$};
%\draw[color=black] (0pt,-10pt) node[right] {\footnotesize $0$};
\clip(-1.5,-7.5) rectangle (17,9.2);
\draw [fill=black,fill opacity=0.5] (8.52,0.98) circle (1.02cm);
\draw [fill=black,fill opacity=0.15] (8.52,0.98) circle (4cm);
\draw [fill=black,fill opacity=0.95] (5.84,3.09) circle (0.21cm);
\draw [fill=black,fill opacity=0.5] (7.65,-3.14) circle (1.32cm);
\draw [line width=2pt] (3.51,0.63) circle (1.9cm);
\draw [dash pattern=on 2pt off 2pt] (3.51,0.63) circle (1.49cm);
\draw (3.51,0.63)-- (0.62,1.33);
\draw (3.51,0.63)-- (6.12,3.39);
\draw (8.52,0.98)-- (3.51,0.63);
\draw (8.52,0.98)-- (14.89,-3.85);
\draw (11.88,-1) node[anchor=north west] {$r_2 +r_1 / ( 1 + \eps$};
\begin{scriptsize}
\fill [color=qqqqff] (3.51,0.63) circle (1.5pt);
\draw[color=qqqqff] (3.71,0.94) node {$q$};
\fill [color=xdxdff] (0.62,1.33) circle (1.5pt);
\draw[color=black] (3.24,1.41) node {$r_1/(1+\eps)$};
\fill [color=qqqqff] (8.52,0.98) circle (1.5pt);
\draw[color=qqqqff] (8.9,1.29) node {$p_2$};
\fill [color=uququq] (6.48,0.83) circle (1.5pt);
\fill [color=qqqqff] (7.83,2.21) circle (1.5pt);
\draw[color=qqqqff] (8.21,2.52) node {$p_3$};
\fill [color=qqqqff] (7.65,-3.14) circle (1.5pt);
\draw[color=qqqqff] (8.01,-2.82) node {$p_4$};
\fill [color=qqqqff] (6.62,0.6) circle (1.5pt);
\draw[color=qqqqff] (6.98,0.92) node {$p_5$};
\fill [color=qqqqff] (1.49,-2.39) circle (1.5pt);
\draw[color=qqqqff] (1.87,-2.08) node {$p_6$};
\fill [color=qqqqff] (1.95,-2.65) circle (1.5pt);
\draw[color=qqqqff] (2.31,-2.33) node {$p_7$};
\fill [color=qqqqff] (2.3,-2.75) circle (1.5pt);
\draw[color=qqqqff] (2.68,-2.43) node {$p_8$};
\fill [color=qqqqff] (2.79,-2.9) circle (1.5pt);
\draw[color=qqqqff] (3.17,-2.57) node {$p_9$};
\fill [color=xdxdff] (14.89,-3.85) circle (1.5pt);
\fill [color=uququq] (5.84,3.09) circle (1.5pt);
\fill [color=xdxdff] (4.52,4.29) circle (1.5pt);
\draw[color=xdxdff] (4.89,4.61) node {$p_1$};
\end{scriptsize}
\end{tikzpicture}
\caption{First two steps of the ANN algorithm. First $p_1$ is 
chosen as the current candidate, and we must compute $\dist(p_2, q)$. After that the algorithm will not compute
distance to any of the points inside the heavily shaded ball or outside the lightly shaded ball that are centered at $p_2$, because their lower bounds allow us to discard them. Note that the point $p_5$, which
is closer to $q$ than $p_1$, also will not be a candidate, and at least one of the points $p_6,p_7,p_8,p_9$ in the annulus between the dashed and solid circle, which are further from $q$ than $p_5$, will be chosen as $c$. This shows that in our algorithm the distance from the candidate to $q$ can drop \textit{slower} than in the bruteforce algorithm, thus Theorem \ref{thm:ann_bound} does not immediately follow from standard backwards analysis. The small black ball between the dashed circle and the solid circle has diameter $v_1 \eps / (1 + \eps)$; it is the ball that we use in the packing argument, because it is smaller than any of the lightly shaded balls that correspond to points like $p_2$ and $p_4$, that is, the points that do not improve $v$.}
\label{fig:ann_illustration}
\end{figure}

>>>>

\documentclass[a4paper,USenglish]{socg-lipics-v2018}
%\documentclass[a4paper,USenglish]{article}

%\usepackage{arxiv}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathbbol}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{url}
\usepackage{floatflt}
\usepackage{mathptmx}
\usepackage{epsfig}
\usepackage{wrapfig}
\usepackage{color}

@@

\definecolor{uququq}{rgb}{0.25,0.25,0.25}
\definecolor{xdxdff}{rgb}{0.49,0.49,1}
\definecolor{qqqqff}{rgb}{0,0,1}
\begin{figure}[ht!]
    \centering
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=0.5cm,y=0.5cm]
%\draw[->,color=black] (-1.5,0) -- (17,0);
%\foreach \x in {,2,4,6,8,10,12,14,16}
%\draw[shift={(\x,0)},color=black] (0pt,2pt) -- (0pt,-2pt) node[below] {\footnotesize $\x$};
%\draw[->,color=black] (0,-7.5) -- (0,9.2);
%\foreach \y in {-6,-4,-2,2,4,6,8}
%\draw[shift={(0,\y)},color=black] (2pt,0pt) -- (-2pt,0pt) node[left] {\footnotesize $\y$};
%\draw[color=black] (0pt,-10pt) node[right] {\footnotesize $0$};
\clip(-1.5,-7.5) rectangle (17,9.2);
\draw [fill=black,fill opacity=0.5] (8.52,0.98) circle (1.02cm);
\draw [fill=black,fill opacity=0.15] (8.52,0.98) circle (4cm);
\draw [fill=black,fill opacity=0.95] (6.12,3.39) circle (0.42cm);
\draw [fill=black,fill opacity=0.5] (7.65,-3.14) circle (1.32cm);
\draw [line width=2pt] (3.51,0.63) circle (1.9cm);
\draw [dash pattern=on 2pt off 2pt] (3.51,0.63) circle (1.49cm);
\draw (3.51,0.63)-- (0.62,1.33);
\draw (3.51,0.63)-- (6.12,3.39);
\draw (8.52,0.98)-- (3.51,0.63);
\draw (8.52,0.98)-- (14.89,-3.85);
\draw (11.88,-1) node[anchor=north west] {$r_2 +r_1 / ( 1 + \eps$};
\begin{scriptsize}
\fill [color=qqqqff] (3.51,0.63) circle (1.5pt);
\draw[color=qqqqff] (3.71,0.94) node {$q$};
\fill [color=xdxdff] (0.62,1.33) circle (1.5pt);
\draw[color=black] (3.24,1.41) node {$r_1/(1+\eps)$};
\fill [color=qqqqff] (8.52,0.98) circle (1.5pt);
\draw[color=qqqqff] (8.9,1.29) node {$p_2$};
\fill [color=uququq] (6.48,0.83) circle (1.5pt);
\fill [color=qqqqff] (7.83,2.21) circle (1.5pt);
\draw[color=qqqqff] (8.21,2.52) node {$p_3$};
\fill [color=qqqqff] (7.65,-3.14) circle (1.5pt);
\draw[color=qqqqff] (8.01,-2.82) node {$p_4$};
\fill [color=qqqqff] (6.62,0.6) circle (1.5pt);
\draw[color=qqqqff] (6.98,0.92) node {$p_5$};
\fill [color=qqqqff] (1.49,-2.39) circle (1.5pt);
\draw[color=qqqqff] (1.87,-2.08) node {$p_6$};
\fill [color=qqqqff] (1.95,-2.65) circle (1.5pt);
\draw[color=qqqqff] (2.31,-2.33) node {$p_7$};
\fill [color=qqqqff] (2.3,-2.75) circle (1.5pt);
\draw[color=qqqqff] (2.68,-2.43) node {$p_8$};
\fill [color=qqqqff] (2.79,-2.9) circle (1.5pt);
\draw[color=qqqqff] (3.17,-2.57) node {$p_9$};
\fill [color=xdxdff] (14.89,-3.85) circle (1.5pt);
\fill [color=uququq] (5.84,3.09) circle (1.5pt);
\fill [color=xdxdff] (4.52,4.29) circle (1.5pt);
\draw[color=xdxdff] (4.89,4.61) node {$p_1$};
\end{scriptsize}
\end{tikzpicture}
\caption{First two steps of the ANN algorithm. First $p_1$ is 
chosen as the current candidate, and we must compute $\dist(p_2, q)$. After that the algorithm will not compute
distance to any of the points inside the heavily shaded ball or outside the lightly shaded ball that are centered at $p_2$, because their lower bounds allow us to discard them. Note that the point $p_5$, which
is closer to $q$ than $p_1$, also will not be a candidate, and at least one of the points $p_6,p_7,p_8,p_9$ in the annulus between the dashed and solid circle, which are further from $q$ than $p_5$, will be chosen as $c$. This shows that in our algorithm the distance from the candidate to $q$ can drop \textit{slower} than in the bruteforce algorithm, thus Theorem \ref{thm:ann_bound} does not immediately follow from standard backwards analysis. The small black ball between the dashed circle and the solid circle has radius $v_1 \eps / (1 + \eps)$; it is the ball that we use in the packing argument, because it is smaller than any of the lightly shaded balls that correspond to points like $p_2$ and $p_4$, that is, the points that do not improve $v$.}
\label{fig:ann_illustration}
\end{figure}
